---
title: "DATA607-Tidying and Transforming Data"
author: "Alexis Mekueko"
date: "9/23/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```



Github Link: https://github.com/asmozo24/Data607_Assignment5.git

Web link: https://rpubs.com/amekueko/669010



## R Packages
```{r load-packages, message=FALSE}


library(tidyverse) #loading all library needed for this assignment
library(openintro)
#head(fastfood)
library(readxl)
library(readr)
library(plyr)
library(dplyr)
#library(dice)
# #library(VennDiagram)
# #library(help = "dice")
library(DBI)
library(dbplyr)
library(data.table)
library(rstudioapi)
#rstudioapi::isAvailable()
library(RJDBC)
library(odbc)
library(RSQLite)
library(rvest)
library(stringr)
library(readtext)
library(ggpubr)
#library(fitdistrplus)
library(ggplot2)
library(moments)
library(qualityTools)
library(normalp)
library(utils)
library(MASS)
library(qqplotr)
#library(DATA606)
library(RCurl)

```

#connecting to my github respository to import data


```{r}

# pulling raw data from my github respository, Data607_Assignment5
# https://github.com/asmozo24/Data607_Assignment5.git

airlines_arrival1 <- read.table(url("https://raw.githubusercontent.com/asmozo24/Data607_Assignment5/master/Arrival_Delays_02_airlines%20.csv"), sep = ",")

airlines_arrival1  ## header not clean
View(airlines_arrival1)

# This access from github to Rstudio also work 
 airlines_arrival2 <- read.csv("https://raw.githubusercontent.com/asmozo24/Data607_Assignment5/master/Arrival_Delays_02_airlines%20.csv",header=TRUE,stringsAsFactors=FALSE)
 airlines_arrival2 # this is to check it is the actual dataframe I am looking for ....this works too.


```

# connection to working directory and import data from local drive

```{r }

setwd("~/R/DATA607_Assg5") #set the working directory to access dataframe
  # display dataframe from local drive
# newDataName <- read.csv("https://raw.githubusercontent.com/asmozo24/Data607_Assignment5/master/Arrival_Delays_02_airlines%20.csv", sep = ",", na.strings = "NA", strip.white = TRUE, stringsAsFactors = FALSE)

Arrival_Delays_02_airlines_ <- read.csv("C:/Users/Petit Mandela/Documents/R/DATA607_Assg5/Arrival_Delays_02_airlines .csv")

Arrival_Delays_02_airlines_

```





### Connecting to SQL server
--this connection works but requires password, I am going to comment this block because preventing the knit process.
 --Now we succeeded in reading Arrival_Delays_02_airlines full table into R dataframe.....

 ```{r connection to sql} #sql connection = con
 
  con <- dbConnect(odbc(),
                   Driver = "SQL Server",
                   Server = "ATM\\ATMSERVER", 
                   Database = "Data607_Airline_Delays_Db",
                   UID = "Alex", 
                   PWD = rstudioapi::askForPassword("Database password"),
                   Port = 1433)
 
  dbListFields(con, "Arrival_Delays_02_airlines")
  airlines_arrival <- dbReadTable(con, "Arrival_Delays_02_airlines") 
  airlines_arrival 


 ```


## Handling missings data
I will use Github link

```{r }

is.na(airlines_arrival2) # checking if there is a missing data in the dataset, return is yes
sum(is.na(airlines_arrival2)) # file to big, checking the sum of all missing data (return is 09 missing data)
airlines_arrival4 <- na.omit(airlines_arrival2) # delete/remove the missings data because it is an imcomplete observation
airlines_arrival4
View(airlines_arrival4)

```


## Tidying data
I will use Github link
What city has the most arrivals all airlines combined? Answer. Phoenix appears to be have most arrivals or the most destination all airlines combined.
Thus, Phoenix is the most visited city and probably collecting more dividends than other cities.

What city has the least arrivals all airlines combined.? Answer: San Diego appears to have the least arrivals or the least destination all airlines combined.
San Diego is the least visited, the mayor needs to work on making the city more attractive. Maybe trying to find out why most people go to Phoenix.

```{r }
row.names(airlines_arrival4) <- NULL
airlines_arrival4

#renaming the no name columns
names(airlines_arrival4)[1] <- 'airlines'
names(airlines_arrival4)[2] <- 'arrival_status'

# Another way of renaming
# airlines_arrival4 <- airlines_arrival4 %>% 
#   rename(airlines = X,  arrival_status = X.1)

airlines_arrival4

view(airlines_arrival4)

#filling  airlines companies
airlines_arrival4$airlines[2] <- c("ALASKA")
airlines_arrival4$airlines[4] <- c("AMWEST")
airlines_arrival4

summary(airlines_arrival4)

# What city has the most arrivals all airlines combined? Answer. Phoenix appears to be have most arrivals or the most destination all airlines combined.
# Thus, Phoenix is the most visited city and probably collecting more dividends than other cities.

# What city has the least arrivals all airlines combined.? Answer: San Diego appears to have the least arrivals or the least destination all airlines combined.
# San Diego is the least visited, the mayor needs to work on making the city more attractive. Maybe trying to find out why most people go to Phoenix.

total_Los_Angeles <- sum(airlines_arrival4$Los.Angeles)     # 1370
total_Phoenix <- sum(airlines_arrival4$Phoenix)             # 5488
total_San.Diego <- sum(airlines_arrival4$San.Diego)         # 680
total_San.Francisco <- sum(airlines_arrival4$San.Francisco) # 1054
total_Seattle <- sum(airlines_arrival4$Seattle)             # 2408

total_arrivals <- c(total_Los_Angeles, total_Phoenix, total_San.Diego, total_San.Francisco, total_Seattle)
cities <- c('Los Angeles', 'Phoenix', 'San Diego', 'San Francisco', 'Seattle')

# Barplot to visualize the number 
Cities_Arrivals <- data.frame(cities, total_arrivals)
View(Cities_Arrivals)
barplot(Cities_Arrivals$total_arrivals, names.arg = Cities_Arrivals$cities, main= "Total Arrivals per City", xlab = "Cities", ylab = "Total Arrivals",  col = rainbow(6))

# another way to do a barplot 
 ggplot(data = Cities_Arrivals, aes(x= cities, y=total_arrivals)) +
    geom_bar(stat="identity", fill = rainbow(5), width = 0.5)

 
```


## Let's look at the data a little bit different

```{r }

#let call gather data in columns

airlines_arrival5 <- airlines_arrival4 %>%
  gather(key = city, value = "total_arrivals", 3:7)


airlines_arrival6 <- airlines_arrival5 %>%
  pivot_wider(names_from = "arrival_status", values_from = "total_arrivals")


# looking at the view (airlines_arrival5), by filtering descending order, I can say Alaska has the least number of delay arrivals which is Phoenix, AMWEST has  the most number of arrivals in time in Phoenix.



```
